---
title: "animal-model-tutorial"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* Goal here is to learn about the animal model, and figure out how you can run it without a pedigree.

* Use sample data from [Pierre de Villemereuil’s tutorial](http://devillemereuil.legtux.org/wp-content/uploads/2012/12/tuto_en.pdf) with and without a pedigree. Compare results.

* Figure out what the differences in the results mean!


```{r, warning = FALSE, message = FALSE}
library(MCMCglmm)
```

Read in data

```{r}
pedigreemulti <- read.table('data-raw/pedigreemulti.txt',header=T)
datamulti <- read.table('data-raw/datamulti.txt',header=T)
models_run <- load("data-processed/modelmulti.RData") ## this has all the outputs from the models run in this script.
```

Glance at the data
```{r}
head(datamulti)
```

```{r}
head(pedigreemulti)
```

```{r}
tail(pedigreemulti)
```


Specify priors for the animal effect and the residual variance

```{r}
prior <- list(R=list(V=diag(2)/2,nu=2),
		 G=list(G1=list(V=diag(2)/2,nu=2)))
```

Set up and run model

```{r,eval=FALSE}
modelmulti <- MCMCglmm(cbind(phen1,phen2)~trait-1, ## the notation trait-1 allows us to estimate the mean of each trait, instead of the contrast between the two traits
					 random=~us(trait):animal, ##  define the structure of the variance-covariance matrix for the random effects 
					 rcov=~us(trait):units,  ##  define the structure of the variance-covariance matrix for the residual variances
					 family=c("gaussian","gaussian"), ## family requires a vector with the data distribution of each trait
					 prior=prior,
					 pedigree=pedigreemulti, ## specifies the pedigree; one row for each individual, each column specifies the parents for each individual
					 data=datamulti, ## includes two traits for each individual, i.e. phen1 and phen 2
					 nitt=100000,
					 burnin=1000,
					 thin=10)

```

Look at the outputs.

```{r}
summary(modelmulti)
```

Plot it. 

```{r}
plot(modelmulti$Sol)
```


Calculate heritabilities

```{r}
herit1<-modelmulti$VCV[,'traitphen1:traitphen1.animal']/ (modelmulti$VCV[,'traitphen1:traitphen1.animal']+ modelmulti$VCV[,'traitphen1:traitphen1.units'])
herit2<-modelmulti$VCV[,'traitphen2:traitphen2.animal']/(modelmulti$VCV[,'traitphen2:traitphen2.animal']+modelmulti$VCV[,'traitphen2:traitphen2.units'])
mean(herit1)
mean(herit2)
```

Calculate genetic correlations among the two traits
```{r}
corr.gen<-modelmulti$VCV[,'traitphen1:traitphen2.animal']/
	sqrt(modelmulti$VCV[,'traitphen1:traitphen1.animal']*modelmulti$VCV[,'traitphen2:traitphen2.animal'])
	mean(corr.gen)
```

Now run the same thing, but just don't give it a pedigree (i.e. remove that argument altogether)

```{r, eval = FALSE}
modelmulti_no_ped <- MCMCglmm(cbind(phen1,phen2)~trait-1, ## the notation trait-1 allows us to estimate the mean of each trait, instead of the contrast between the two traits
					   random=~us(trait):animal, ##  define the structure of the variance-covariance matrix for the random effects 
					   rcov=~us(trait):units,  ##  define the structure of the variance-covariance matrix for the residual variances
					   family=c("gaussian","gaussian"), ## family requires a vector with the data distribution of each trait
					   prior=prior,
					   # pedigree=pedigreemulti, ## specifies the pedigree; one row for each individual, each column specifies the parents for each individual
					   data=datamulti, ## includes two traits for each individual, i.e. phen1 and phen 2
					   nitt=100000,
					   burnin=1000,
					   thin=10)
```


Look at output
```{r}
summary(modelmulti_no_ped)
```

Plot it. 

```{r}
plot(modelmulti_no_ped$Sol)
```


Calculate heritabilities

```{r}
herit1_no_ped<-modelmulti_no_ped$VCV[,'traitphen1:traitphen1.animal']/ (modelmulti_no_ped$VCV[,'traitphen1:traitphen1.animal']+ modelmulti_no_ped$VCV[,'traitphen1:traitphen1.units'])
herit2_no_ped<-modelmulti_no_ped$VCV[,'traitphen2:traitphen2.animal']/(modelmulti_no_ped$VCV[,'traitphen2:traitphen2.animal']+modelmulti_no_ped$VCV[,'traitphen2:traitphen2.units'])
mean(herit1_no_ped)
mean(herit2_no_ped)
```

Calculate genetic correlations among the two traits. 

```{r}
corr.gen_no_ped<-modelmulti_no_ped$VCV[,'traitphen1:traitphen2.animal']/
	sqrt(modelmulti_no_ped$VCV[,'traitphen1:traitphen1.animal']*modelmulti_no_ped$VCV[,'traitphen2:traitphen2.animal'])
mean(corr.gen_no_ped)
```

